---
title: 'ADA on Employment: Analysis'
author: "Young Ahn"
date: "2025-01-21"
output:
  pdf_document: default
  html_document: default
---

Analysis using SIPP 1990 Panel with three analysis configurations:

1. **Monthly Analysis by Rotation** - All 32 months per rotation
   - time_index_monthly: relative to rotation (1 = first month in rotation group)
   - NOT calendar-based; each rotation starts at its own first reference month
   - Treatment: July 1990 (ADA signing), which corresponds to:
     - Rotation 1: time_index = 7 (first reference month = Jan 1990)
     - Rotation 2: time_index = 10 (first reference month = Oct 1989)
     - Rotation 3: time_index = 9 (first reference month = Nov 1989)
     - Rotation 4: time_index = 8 (first reference month = Dec 1989)
   - Each rotation analyzed separately with rotation-specific treatment timing

2. **Wave-based Analysis by Rotation** - 8 waves (first month of each)
   - time_index_wave = interview_wave (1-8)
   - Treatment: Wave 3
   - Each rotation analyzed separately

3. **Combined Analysis (Jan/May/Sep)** - All 4 rotations combined
   - time_index_combined: 1 = Jan 1990, 2 = May 1990, 3 = Sep 1990, ..., 8 = May 1992 (every 4 months)
   - Treatment: Sep 1990 (first post-ADA period)
   - 2 pretreatment + varying post-treatment periods (t4-t8)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf', fig.width = 8, fig.height = 5)

library(did)
library(dplyr)
library(haven)
library(ak)
library(ggpubr)
library(RColorBrewer)
library(tidyr)
```

```{r setup-parameters}
# Data path
data_ada_path <- "data/generated/ada.RData"

# State and outcome information data directories
outcome_pairs_dir <- "data/empirical-specs/empirical_specs.csv"

# Estimation options
opts = list(se_method = "boot", bootstrap_counts = 500,
            multistart_mc_core_counts_short_run = 10,
            multistart_mc_core_counts_long_run = 4,
            bootstrap_mc_core_counts = 4)

y_lab_function_transition <- function(y_name_from, y_name_to) {
  y_name_to <- ifelse(y_name_to == "olf", "OLF",
                        ifelse(y_name_to == "employed", "Employed", "Unemployed"))
  y_name_from <- ifelse(y_name_from == "olf", "OLF",
                        ifelse(y_name_from == "employed", "Employed", "Unemployed"))

  sprintf("%s-to-%s Rates", y_name_from, y_name_to)
}
```

# Load data
```{r load-data}
# Check if file exists first, and load
if (!file.exists(data_ada_path)) {
  stop(paste0("File does not exist: ", data_ada_path,
              "; run ada-on-employment-data-prep.Rmd first."))
}

load(data_ada_path)

cat("=== Data Summary ===\n")
cat(paste("Number of total observations:", nrow(event_study_df), "\n"))
cat(paste("Number of unique units:", length(unique(event_study_df$id)), "\n"))

# Show sample sizes for each analysis type
cat("\n=== Sample Sizes by Analysis Configuration ===\n")

cat("\nCombined Analysis (all rotations, Jan/May/Sep):\n")
for (t in 4:8) {
  sample_col <- paste0("sample_combined_t", t)
  n_units <- length(unique(event_study_df$id[event_study_df[[sample_col]] == TRUE]))
  cat(sprintf("  t%d (%d periods): %d units\n", t, t, n_units))
}

cat("\nWave-based Analysis by Rotation (8 waves each):\n")
for (r in 1:4) {
  sample_col <- paste0("sample_wave_rot_", r)
  n_units <- length(unique(event_study_df$id[event_study_df[[sample_col]] == TRUE]))
  cat(sprintf("  Rotation %d: %d units\n", r, n_units))
}

cat("\nMonthly Analysis by Rotation (32 months each):\n")
for (r in 1:4) {
  sample_col <- paste0("sample_monthly_rot_", r)
  n_units <- length(unique(event_study_df$id[event_study_df[[sample_col]] == TRUE]))
  cat(sprintf("  Rotation %d: %d units\n", r, n_units))
}

# Show calendar month mapping for combined analysis
cat("\n=== Calendar Period Mapping (Combined Analysis) ===\n")
calendar_periods <- event_study_df %>%
  filter(!is.na(time_index_combined)) %>%
  select(time_index_combined, calendar_year, calendar_month) %>%
  distinct() %>%
  arrange(time_index_combined)

for (i in 1:nrow(calendar_periods)) {
  month_name <- month.abb[calendar_periods$calendar_month[i]]
  pre_post <- ifelse(calendar_periods$time_index_combined[i] < 3, "(pretreatment)",
                     ifelse(calendar_periods$time_index_combined[i] == 3,
                            "(first post-treatment)", "(post-treatment)"))
  cat(sprintf("  %d = %s %d %s\n",
              calendar_periods$time_index_combined[i],
              month_name,
              calendar_periods$calendar_year[i],
              pre_post))
}
```

# Analysis
```{r load-empirical-specs}
# Load empirical specifications
empirical_specs <- read.csv(outcome_pairs_dir, row.names = NULL) %>%
  filter(included_in_analysis == 1) %>%
  empirical_specs_df_to_empirical_specs()

cat("Number of empirical specifications:", length(empirical_specs), "\n")
for (i in seq_along(empirical_specs)) {
  cat(paste0("  ", i, ". ", empirical_specs[[i]]$sample_name, "\n"))
}
```

```{r run-analysis}
empirical_analysis_results <- lapply(empirical_specs, perform_empirical_analysis,
                                      event_study_df, opts = opts)
model_pairs <- lapply(empirical_analysis_results, "[[", "transition_models")
models <- unlist(model_pairs, recursive = FALSE)

cat("\nAnalysis complete. Number of models:", length(models), "\n")
```

```{r plot-averages}
plot_average_pairs <- lapply(models, plot_averages)

plot_average_pairs
```


## Aggregated Transition Plots (3x3 Grid)

Helper function to create aggregated 3x3 transition plot from individual plots:
```{r helper-aggregate-transitions}
#' Create aggregated 3x3 transition plot
#' @param transition_plots List of individual transition plots from plot_transitions()
#' @param y_names Vector of outcome names (e.g., c("employed", "unemployed", "olf"))
#' @return A ggplot object with all 9 transitions arranged in a 3x3 grid
create_aggregated_transition_plot <- function(transition_plots, y_names) {
  # Create plotlist in row-major order (rows = y_to, cols = y_from)
  plotlist <- unlist(
    lapply(y_names, function(y_to) {
      lapply(y_names, function(y_from) {
        plot_name <- paste0("-from-", y_from, "-to-", y_to)
        transition_plots[[plot_name]]
      })
    }),
    recursive = FALSE
  )

  # Arrange in grid
 ggpubr::ggarrange(
    plotlist = plotlist,
    ncol = length(y_names),
    nrow = length(y_names),
    common.legend = TRUE,
    legend = "bottom"
  )
}
```

```{r plot-transitions-aggregated-uniform-ci-pretreatment, fig.width=12, fig.height=12}
y_names <- c("employed", "unemployed", "olf")

# Pretreatment only
plot_transition_list_pre <- lapply(models, plot_transitions, uniform_ci = TRUE,
                                    pretreatment_only = TRUE,
                                    y_lab_function = y_lab_function_transition)

for (i in seq_along(plot_transition_list_pre)) {
  model_name <- names(plot_transition_list_pre)[i]
  if (is.null(model_name)) model_name <- paste("Model", i)
  cat("\n###", model_name, "(Pretreatment only, Uniform CI)\n")
  p <- create_aggregated_transition_plot(plot_transition_list_pre[[i]], y_names)
  print(p)
}
```

All outcomes:
```{r plot-trends-all-outcomes}
for (model in models) {
  model_copy <- model
  model_copy$event_study_df <- model_copy$event_study_df %>%
    mutate(labor_force_status = ifelse(labor_force_status == "olf", "OLF", labor_force_status)) %>%
    mutate(labor_force_status = ifelse(labor_force_status == "employed", "Employed", labor_force_status)) %>%
    mutate(labor_force_status = ifelse(labor_force_status == "unemployed", "Unemployed", labor_force_status))
  model_copy$Y$names <- c("Employed", "Unemployed", "OLF")
  print(plot_data_trends_all_outcome(model_copy))
}
```


```{r plot-atts}
plot_atts <- lapply(empirical_analysis_results, plot_all_att_comparison)

plot_atts
```

Simplified ATT comparison (DiD vs Our Model):
```{r plot-atts-simple}
plot_atts_simple <- lapply(empirical_analysis_results, plot_att_comparison_simple)

plot_atts_simple
```


# Decomposing ATT
## Draw plots
```{r decomposition-plots}
# Get all outcome_varnames
outcome_varnames <- sapply(models, `[[`, "outcome_varname")

# Find the index of the first occurrence of each unique outcome_varname
first_indices <- match(unique(outcome_varnames), outcome_varnames)

# Extract one model per unique outcome_varname
representative_models <- models[first_indices]

# Label function for decomposition plot legend
outcome_label_function <- function(x) {
  ifelse(x == "olf", "OLF",
         ifelse(x == "employed", "Employed",
                ifelse(x == "unemployed", "Unemployed", x)))
}

# Generate decomposition plots for each model
lapply(representative_models, generate_all_decomposition_plots,
       outcome_label_function = outcome_label_function)
```
