---
title: 'ADA on Employment: Data Preparation'
author: "Young Ahn"
date: "2025-04-15"
output:
  pdf_document: default
  html_document: default
---

SIPP 1990 Panel timing (Table 1, page 12 in raw/sipp1990fp.pdf)
- Rotation 1: First interview May 1990, first reference month Jan 1990
- Rotation 2: First interview Feb 1990, first reference month Oct 1989
- Rotation 3: First interview Mar 1990, first reference month Nov 1989
- Rotation 4: First interview Apr 1990, first reference month Dec 1989
The panel runs for 32 months (8 waves x 4 months each)

Three analysis configurations:

1. **Monthly Analysis by Rotation** - All 32 months per rotation
   - time_index_monthly: relative to rotation (1 = first month in rotation group)
   - NOT calendar-based; each rotation starts at its own first reference month
   - Treatment: July 1990 (ADA signing), which corresponds to:
     - Rotation 1: time_index = 7 (first reference month = Jan 1990)
     - Rotation 2: time_index = 10 (first reference month = Oct 1989)
     - Rotation 3: time_index = 9 (first reference month = Nov 1989)
     - Rotation 4: time_index = 8 (first reference month = Dec 1989)
   - Each rotation analyzed separately with rotation-specific treatment timing

2. **Wave-based Analysis by Rotation** - First month of each wave (8 waves)
   - time_index_wave = interview_wave (1-8)
   - Treatment period: Wave 3 (time_index_wave = 3)
   - Each rotation analyzed separately

3. **Combined Analysis (Jan/May/Sep)** - All 4 rotations combined
   - time_index_combined: 1 = Jan 1990, 2 = May 1990, 3 = Sep 1990, ..., 8 = May 1992 (every 4 months)
   - Treatment period: time_index = 3 (Sep 1990, first post-treatment)
   - All rotation groups combined


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf', fig.width = 8, fig.height = 5)

library(did)
library(dplyr)
library(haven)
library(ak)
library(ggpubr)
library(RColorBrewer)
library(tidyr)
```

```{r setup-parameters}
extraction_dir <- "data/extracted/"
zip_file_transition <- c("data/raw/sip90fp.zip", "data/raw/sip90t3.zip")
data_final_dir <- "data/extracted/"
data_final_dir_w1 <- file.path(data_final_dir, "sip90fp.dta")
data_topical_dir <- file.path(data_final_dir, "sip90t3.dta")
data_ada_path <- "data/generated/ada.RData"

# State and outcome information data directories
outcome_pairs_dir <- "data/empirical-specs/empirical_specs.csv"

# Prefixes used for columns for Monthly data 
monthly_data_columns_prefix <- c("esr_", "pp_ear", "age_")

# Age range
age_min <- 21
age_max <- 58

# Subsample counts (actual data contain more than 20,000 individuals)
# Use Inf to include all individuals
N_SUBSAMPLE <- Inf

# Time index parameters
# ADA was signed July 26, 1990
#
# Three analysis configurations:
#
# 1. Monthly Analysis: All 32 months per rotation
#    time_index_monthly: relative to rotation (1 = first month)
#    NOT calendar-based; each rotation starts at its own first reference month
#    Treatment: July 1990 (ADA signing) - rotation-specific time indices:
#      - Rotation 1: time_index = 7 (first reference month = Jan 1990)
#      - Rotation 2: time_index = 10 (first reference month = Oct 1989)
#      - Rotation 3: time_index = 9 (first reference month = Nov 1989)
#      - Rotation 4: time_index = 8 (first reference month = Dec 1989)
#
# 2. Wave-based Analysis: First month of each wave (8 waves)
#    time_index_wave = interview_wave (1-8)
#    Treatment: wave 3
#
# 3. Combined Analysis (Jan/May/Sep): All rotations combined
#    time_index_combined: 1 = Jan 1990, 2 = May 1990, 3 = Sep 1990, ..., 8 = May 1992
#    Treatment: time_index = 3 (September 1990)
#
TIME_INDEX_TREATED_MONTHLY <- 7  # July 1990 (when ADA was signed) - for rotation 1 only
# Rotation-specific treatment timing (July 1990 at correct index for each rotation)
# Each rotation starts at a different calendar month, so July 1990 corresponds to:
TIME_INDEX_TREATED_MONTHLY_ROT_1 <- 7   # Jan 1990 + 6 = July 1990
TIME_INDEX_TREATED_MONTHLY_ROT_2 <- 10  # Oct 1989 + 9 = July 1990
TIME_INDEX_TREATED_MONTHLY_ROT_3 <- 9   # Nov 1989 + 8 = July 1990
TIME_INDEX_TREATED_MONTHLY_ROT_4 <- 8   # Dec 1989 + 7 = July 1990
TIME_INDEX_TREATED_WAVE <- 3     # Wave 3
TIME_INDEX_TREATED_COMBINED <- 3 # September 1990 (first post-treatment period)

# Legacy variable for backward compatibility
TIME_INDEX_TREATED <- TIME_INDEX_TREATED_COMBINED

```

```{r load-data}
#' Unzip Files to a Directory
#'
#' This function unzips the files contained in the specified zip file and saves them in the specified extraction directory.
#'
#' @param zip_file_transition A character string representing the transition to the zip file. Defaults to "data/raw/jtpa_national_evaluation.zip".
#' @param extraction_dir A character string representing the directory where the files should be extracted. Defaults to "data/extracted/jtpa_national_evaluation".
#'
#' @return This function does not return a value. It unzips the specified zip file into the specified extraction directory.
#' @examples
#' unzip_to_dir()
#' unzip_to_dir("transition/to/your.zip", "transition/to/extraction_dir")
unzip_to_dir <- function(zip_file_transition = "data/raw/jtpa_national_evaluation.zip",
                         extraction_dir = "data/extracted/jtpa_national_evaluation") {

  # Check if the extraction directory exists; if not, create it
  if (!dir.exists(extraction_dir)) {
    dir.create(extraction_dir)
  }

  # Unzip the files
  sapply(zip_file_transition, function(x) unzip(zipfile = x, exdir = extraction_dir))
}

# Uncomment the following lines if ran for the first time
unzip_to_dir(zip_file_transition = zip_file_transition, extraction_dir = extraction_dir)
```


Data cleaning:

```{r}
# Load the state codes
# state_codes <- read.csv(tst_intv_codebook_dir)
# aca_status_df <- read.csv(aca_status_dir)
empirical_specs_df <- read.csv(outcome_pairs_dir, row.names = NULL) %>%
  filter(included_in_analysis == 1)

# Generate a `treated` variable based on ACA initial expansion status. (`state_to_treated`)
# Could be NaN if out of state or foreign coutnries
disab_to_treated <- function(df) {
  df %>%
    mutate(treated = ifelse(state %in% aca_status_df$state[aca_status_df$aca_introduced_jan2014 == 1], 1, 0)) %>%
    mutate(treated = ifelse(state %in% aca_status_df$state[aca_status_df$aca_already_implmented_jan2014 == 1], NaN, treated)) %>%
    mutate(treated = ifelse(state %in% aca_status_df$state, treated, NaN))
}

filter_by_age <- function(df_long, age_min = -Inf, age_max = Inf) {
  # Remove units (id) whose ages are outside the specified range in any month
  df_long %>%
    group_by(id) %>%
    # Check if ALL age values are within the bounds
    filter(all(age >= age_min & age <= age_max)) %>%
    ungroup()
}

add_employed <- function(df) {
  df %>% 
    mutate(employed = ifelse(esr > 0 & esr < 6, 1, 0)) %>%
    mutate(unemployed = ifelse(esr >= 6 & esr <= 7, 1, 0)) %>%
    mutate(olf = !(employed | unemployed)) %>%
    mutate(labor_force_status = ifelse(employed == 1, "employed",
                                        ifelse(unemployed == 1, "unemployed", "olf")))
}

filter_complete_data <- function(df_merged) {
  # Determine the full range of month_index
  full_time_index <- seq(min(df_merged$time_index), max(df_merged$time_index))
  
  # Group by id and check for complete month_index
  df_complete_check <- df_merged %>%
    group_by(id) %>%
    summarize(has_complete_monthly_data = all(full_time_index %in% time_index),
              has_complete_esr_data = !any(esr == 0)) %>%
    ungroup() %>%
    mutate(has_complete_monthly_data = ifelse(is.na(has_complete_monthly_data), FALSE, has_complete_monthly_data)) %>%
    mutate(has_complete_data = has_complete_monthly_data & has_complete_esr_data) 
  
  # Merge the result back to df and filter based on has_complete_data
  event_study_df <- df_merged %>%
    left_join(df_complete_check, by = "id") %>%
    filter(has_complete_data == TRUE)
  
  return(event_study_df)
}

wide_to_long_df_sipp1990 <- function(wide_df, monthly_data_columns_prefix) {
  # Build a regex pattern to match the prefixes
  prefix_pattern <- paste0("^(", paste(monthly_data_columns_prefix, collapse = "|"), ")\\d+")

  # SIPP 1990 Panel timing (Table 1, page 12 in raw/sipp1990fp.pdf)
  # - Rotation 1: First interview May 1990, first reference month Jan 1990
  # - Rotation 2: First interview Feb 1990, first reference month Oct 1989
  # - Rotation 3: First interview Mar 1990, first reference month Nov 1989
  # - Rotation 4: First interview Apr 1990, first reference month Dec 1989
  # The panel runs for 32 months (8 waves x 4 months each)

  # Reshape the data from wide to long format
  long_df <- wide_df %>%
    select(suseqnum, rot, disab, disab_nwr, disab_any, white, male, state,
           # Monthly analysis treatment cohorts
           treatment_cohort_monthly_disab, treatment_cohort_monthly_disab_any,
           # Rotation-specific monthly treatment cohorts (July 1990 at correct index for each rotation)
           treatment_cohort_monthly_rot_1_disab, treatment_cohort_monthly_rot_2_disab,
           treatment_cohort_monthly_rot_3_disab, treatment_cohort_monthly_rot_4_disab,
           # Wave-based analysis treatment cohorts
           treatment_cohort_wave_disab, treatment_cohort_wave_disab_any,
           # Combined analysis treatment cohorts
           treatment_cohort_combined_disab, treatment_cohort_combined_disab_any,
           # Legacy treatment cohorts
           treatment_cohort_disab, treatment_cohort_disab_any,
           higrade1, higrade2, higrade3, higrade4,
           higrade5, higrade6, higrade7, higrade8,
           matches(prefix_pattern)) %>%
    tidyr::pivot_longer(
      cols = matches(prefix_pattern),
      names_to = c(".value", "interview_month"),
      names_pattern = paste0("(", paste(monthly_data_columns_prefix, collapse = "|"), ")(\\d+)")
    ) %>%
    rename_with(~ stringr::str_remove(., "_$"))  %>%
    mutate(interview_month = as.integer(interview_month)) %>%
    rename(id = suseqnum) %>%
    mutate(
        rot_offset = case_when(
          rot == 1 ~ 0L,
          rot == 2 ~ -3L,
          rot == 3 ~ -2L,
          rot == 4 ~ -1L
        ),
        # time_index: 1 = Jan 1990, 2 = Feb 1990, ...,
        # 0 = Dec 1989, -1 = Nov 1989, -2 = Oct 1989
        time_index = interview_month + rot_offset,
        # Convert to calendar year and month
        calendar_year = 1990L + (time_index - 1L) %/% 12L,
        calendar_month = ((time_index - 1L) %% 12L) + 1L
      ) %>%
    # Compute interview wave (1-8, each wave covers 4 months)
    mutate(interview_wave = (interview_month - 1L) %/% 4L + 1L) %>%
    mutate(treated = (disab == 1),
           treatment_cohort = treated * TIME_INDEX_TREATED,
           treatment_cohort_nwr = disab_nwr * TIME_INDEX_TREATED) %>%
    mutate(sample_calendar_month = (time_index > 0) & (time_index <= max(interview_month) - 3))

  long_df
}

# Adjust panel df to event study df
panel_df_to_event_study_df <- function(df_merged, id_col = "id", index_col = "time_index", 
                                       treated_col = "treated") {
  df_merged %>%
    rename(id = !!sym(id_col), time_index = !!sym(index_col), treated = !!sym(treated_col)) %>%
    mutate(
      treatment_cohort = treated * TIME_INDEX_TREATED
    )
}

subsample <- function(df, N, idname = "id") {
  # Subsample the same population for reproducibility
  set.seed(1234)
  
  # Get unique IDs from the dataframe
  unique_ids <- unique(df[[idname]])
  
  # Check if N is greater than the number of unique IDs
  if (N > length(unique_ids)) {
    warning("N is larger than the number of unique IDs in the dataframe.")
  }
  
  # Randomly sample N IDs
  sampled_ids <- sample(unique_ids, min(length(unique_ids), N))
  
  # Subset the dataframe to include only the rows with the sampled IDs
  df_subsample <- df %>% filter(.data[[idname]] %in% sampled_ids)
  
  return(df_subsample)
}

expand_column <- function(df, column_name = "rot", idname = "id",
                          fill_value = TRUE) {
  expanded <- df %>%
    select(!!sym(idname), !!sym(column_name)) %>%
    group_by(!!sym(idname)) %>%
    summarize(!!column_name := first(!!sym(column_name))) %>%
    mutate(row_id = row_number(), value = fill_value) %>%
    pivot_wider(
        names_from = !!sym(column_name),
        values_from = value,
        values_fill = FALSE,
        names_prefix = paste0(column_name, "_")
    ) %>%
    select(-row_id) 
  df %>% 
    left_join(expanded, by = idname)
}

add_wave <- function(long_df, months_per_wave = 4) {
  long_df %>%
    # sample_wave indicates first month of each wave (for sampling purposes)
    mutate(sample_wave = 1*((interview_month %% months_per_wave) == 1))
    # Note: interview_wave is now computed in wide_to_long_df_sipp1990()
}
```

```{r data-topical}
df_topical_raw <- haven::read_dta(data_topical_dir)

disab_cond_1_colnames <- c("tm8926", "tm8928")
disab_cond_2_colnames <- c("tm8838", "tm8842", "tm8846", "tm8850", "tm8854", 
                           "tm8858", "tm8862", "tm8866", "tm8870", "tm8871")
disab_cond_3_colnames <- c("tm8802", "tm8804", "tm8810", "tm8814", "tm8818", 
                           "tm8822", "tm8826", "tm8830", "tm8834")
disab_cond_4_colnames <- c("tm8902", "tm8904", "tm8906", "tm8908", "tm8910")


df_topical <- df_topical_raw %>%
  mutate(disab_cond_1 = rowSums(across(all_of(disab_cond_1_colnames), ~ . == 1), na.rm = TRUE)) %>%
  mutate(disab_cond_2 = rowSums(across(all_of(disab_cond_2_colnames), ~ . == 1), na.rm = TRUE)) %>%
  mutate(disab_cond_3 = rowSums(across(all_of(disab_cond_3_colnames), ~ . == 1), na.rm = TRUE)) %>%
  mutate(disab_cond_4 = rowSums(across(all_of(disab_cond_4_colnames), ~ . == 1), na.rm = TRUE)) %>%
  mutate(disab_nwr_satisfied = (disab_cond_1 > 0) | (disab_cond_2 > 1) | (disab_cond_3 > 0) | (disab_cond_4 > 0)) %>%
  rename(pp_id = id, pp_entry = entry, pp_pnum = pnum) %>%
  select(pp_id, pp_entry, pp_pnum, disab_nwr_satisfied, disab_cond_1, disab_cond_2, disab_cond_3, disab_cond_4)

```

```{r data-loading}
df_raw <- haven::read_dta(data_final_dir_w1) 

# some units do not appear due to death or not being born yet
df_raw_merged_with_topical <- df_raw %>%
  mutate(pp_id = as.integer(pp_id)) %>%
  left_join(df_topical, by = c("pp_id", "pp_entry", "pp_pnum")) %>%
  mutate(disab_nwr = ifelse(is.na(disab_nwr_satisfied), 0, (1-disab) * disab_nwr_satisfied)) %>%
  mutate(disab_any = disab_nwr | disab,
         # Treatment cohort columns for different analysis types
         # Monthly: treatment at July 1990 (time_index_monthly = 7 for rotation 1)
         # Legacy variable - uses rotation 1 timing for all rotations (incorrect for rotations 2-4)
         treatment_cohort_monthly_disab = disab * TIME_INDEX_TREATED_MONTHLY,
         treatment_cohort_monthly_disab_any = disab_any * TIME_INDEX_TREATED_MONTHLY,
         # Rotation-specific monthly treatment cohorts (July 1990 at correct index for each rotation)
         treatment_cohort_monthly_rot_1_disab = if_else(rot == 1, as.integer(disab * TIME_INDEX_TREATED_MONTHLY_ROT_1), 0L),
         treatment_cohort_monthly_rot_2_disab = if_else(rot == 2, as.integer(disab * TIME_INDEX_TREATED_MONTHLY_ROT_2), 0L),
         treatment_cohort_monthly_rot_3_disab = if_else(rot == 3, as.integer(disab * TIME_INDEX_TREATED_MONTHLY_ROT_3), 0L),
         treatment_cohort_monthly_rot_4_disab = if_else(rot == 4, as.integer(disab * TIME_INDEX_TREATED_MONTHLY_ROT_4), 0L),
         # Wave-based: treatment at wave 3
         treatment_cohort_wave_disab = disab * TIME_INDEX_TREATED_WAVE,
         treatment_cohort_wave_disab_any = disab_any * TIME_INDEX_TREATED_WAVE,
         # Combined (Jan/May/Sep): treatment at Sep 1990 (time_index_combined = 3)
         treatment_cohort_combined_disab = disab * TIME_INDEX_TREATED_COMBINED,
         treatment_cohort_combined_disab_any = disab_any * TIME_INDEX_TREATED_COMBINED,
         # Legacy columns for backward compatibility
         treatment_cohort_disab = treatment_cohort_combined_disab,
         treatment_cohort_disab_any = treatment_cohort_combined_disab_any,
         white = (race == 1),
         male = (sex == 1),
         state = geo_ste1)
```

```{r data-merging-and-cleaning}
# Process all data - do NOT filter to specific rotations or months early
# All three analysis types (Monthly, Wave, Combined) will be supported from this dataset
event_study_df_all <- df_raw_merged_with_topical %>%
  wide_to_long_df_sipp1990(monthly_data_columns_prefix) %>%
  add_employed() %>%
  filter_by_age(age_min = age_min, age_max = age_max) %>%
  add_wave(months_per_wave = 4)

# Add rotation indicator columns
event_study_df_all <- event_study_df_all %>%
  expand_column("rot", "id") %>%
  drop_na() %>%
  as.data.frame()

# Subsample if needed (for testing)
event_study_df_all <- event_study_df_all %>%
  subsample(N_SUBSAMPLE)

n_units_all <- length(unique(event_study_df_all$id))
print(paste("Total # of units after initial processing:", n_units_all))

# Create time index columns for each analysis type
event_study_df_all <- event_study_df_all %>%
  mutate(
    # Monthly: relative to rotation (1 = first month in rotation, 32 = last month)
    # Each rotation has indices 1-32; treatment at index 7 means 7th month in each rotation
    time_index_monthly = interview_month,

    # Wave-based: use interview_wave (1-8)
    time_index_wave = interview_wave,

    # Combined (Jan/May/Sep): remap to 1-8 for Jan/May/Sep months
    time_index_combined = case_when(
      calendar_year == 1990 & calendar_month == 1 ~ 1L,
      calendar_year == 1990 & calendar_month == 5 ~ 2L,
      calendar_year == 1990 & calendar_month == 9 ~ 3L,
      calendar_year == 1991 & calendar_month == 1 ~ 4L,
      calendar_year == 1991 & calendar_month == 5 ~ 5L,
      calendar_year == 1991 & calendar_month == 9 ~ 6L,
      calendar_year == 1992 & calendar_month == 1 ~ 7L,
      calendar_year == 1992 & calendar_month == 5 ~ 8L,
      TRUE ~ NA_integer_
    )
  )

# Create sample indicators for Monthly Analysis by Rotation
# Monthly analysis uses ALL 32 months per rotation
# Each rotation has time_index_monthly = 1 to 32 (rotation-relative)
event_study_df_all <- event_study_df_all %>%
  mutate(
    # All rotations: time_index_monthly = 1-32
    sample_monthly_rot_1 = rot_1,
    sample_monthly_rot_2 = rot_2,
    sample_monthly_rot_3 = rot_3,
    sample_monthly_rot_4 = rot_4
  )

# Create sample indicators for Monthly Rotation 1 with different post-treatment periods
# Treatment is at month 7 for Rotation 1 (July 1990)
# Post-treatment periods: 12, 14, 20, 22 months (ending at months 19, 21, 27, 29)
event_study_df_all <- event_study_df_all %>%
  mutate(
    sample_monthly_rot_1_post12 = rot_1 & (time_index_monthly <= 19),  # 12 months post-treatment
    sample_monthly_rot_1_post14 = rot_1 & (time_index_monthly <= 21),  # 14 months post-treatment
    sample_monthly_rot_1_post20 = rot_1 & (time_index_monthly <= 27),  # 20 months post-treatment
    sample_monthly_rot_1_post22 = rot_1 & (time_index_monthly <= 29)   # 22 months post-treatment
  )

# Create sample indicators for Monthly Rotation 2 with different post-treatment periods
# Treatment is at month 10 for Rotation 2 (July 1990)
# Post-treatment periods: 12, 14, 20, 22 months (ending at months 22, 24, 30, 32)
event_study_df_all <- event_study_df_all %>%
  mutate(
    sample_monthly_rot_2_post12 = rot_2 & (time_index_monthly <= 22),  # 12 months post-treatment
    sample_monthly_rot_2_post14 = rot_2 & (time_index_monthly <= 24),  # 14 months post-treatment
    sample_monthly_rot_2_post20 = rot_2 & (time_index_monthly <= 30),  # 20 months post-treatment
    sample_monthly_rot_2_post22 = rot_2                                # 22 months post-treatment (all 32 months)
  )

# Create sample indicators for Wave-based Analysis by Rotation
# Wave analysis uses first month of each wave (8 waves total)
event_study_df_all <- event_study_df_all %>%
  mutate(
    # sample_wave = 1 for first month of each wave (already computed in add_wave())
    sample_wave_rot_1 = rot_1 & (sample_wave == 1),
    sample_wave_rot_2 = rot_2 & (sample_wave == 1),
    sample_wave_rot_3 = rot_3 & (sample_wave == 1),
    sample_wave_rot_4 = rot_4 & (sample_wave == 1)
  )

# Create sample indicators for Combined Analysis (all rotations, Jan/May/Sep)
# Combined analysis uses Jan/May/Sep for ALL rotation groups
event_study_df_all <- event_study_df_all %>%
  mutate(
    # Base sample: rows with valid time_index_combined (Jan/May/Sep months)
    sample_combined_base = !is.na(time_index_combined),

    # Different time period configurations (2 pre-treatment + varying post-treatment)
    sample_combined_t4 = sample_combined_base & (time_index_combined <= 4),
    sample_combined_t5 = sample_combined_base & (time_index_combined <= 5),
    sample_combined_t6 = sample_combined_base & (time_index_combined <= 6),
    sample_combined_t7 = sample_combined_base & (time_index_combined <= 7),
    sample_combined_t8 = sample_combined_base  # All 8 periods
  )

# Now filter to units with complete panel for each analysis type
# We need to identify which units have complete data for each configuration

# For Combined analysis: find units with all required time periods
create_complete_panel_indicator <- function(df, sample_col, time_col, max_time) {
  # Get units with complete data for this configuration
  complete_units <- df %>%
    filter(!!sym(sample_col)) %>%
    group_by(id) %>%
    summarize(
      n_periods = n_distinct(!!sym(time_col)),
      has_complete = n_periods == max_time,
      .groups = "drop"
    ) %>%
    filter(has_complete) %>%
    pull(id)

  return(complete_units)
}

# Identify units with complete panels for Combined analysis (t4 through t8)
complete_units_combined_t4 <- create_complete_panel_indicator(event_study_df_all, "sample_combined_t4", "time_index_combined", 4)
complete_units_combined_t5 <- create_complete_panel_indicator(event_study_df_all, "sample_combined_t5", "time_index_combined", 5)
complete_units_combined_t6 <- create_complete_panel_indicator(event_study_df_all, "sample_combined_t6", "time_index_combined", 6)
complete_units_combined_t7 <- create_complete_panel_indicator(event_study_df_all, "sample_combined_t7", "time_index_combined", 7)
complete_units_combined_t8 <- create_complete_panel_indicator(event_study_df_all, "sample_combined_t8", "time_index_combined", 8)

# Identify units with complete panels for Wave analysis (8 waves)
complete_units_wave_rot_1 <- create_complete_panel_indicator(event_study_df_all, "sample_wave_rot_1", "time_index_wave", 8)
complete_units_wave_rot_2 <- create_complete_panel_indicator(event_study_df_all, "sample_wave_rot_2", "time_index_wave", 8)
complete_units_wave_rot_3 <- create_complete_panel_indicator(event_study_df_all, "sample_wave_rot_3", "time_index_wave", 8)
complete_units_wave_rot_4 <- create_complete_panel_indicator(event_study_df_all, "sample_wave_rot_4", "time_index_wave", 8)

# Identify units with complete panels for Monthly analysis (32 months)
complete_units_monthly_rot_1 <- create_complete_panel_indicator(event_study_df_all, "sample_monthly_rot_1", "time_index_monthly", 32)
complete_units_monthly_rot_2 <- create_complete_panel_indicator(event_study_df_all, "sample_monthly_rot_2", "time_index_monthly", 32)
complete_units_monthly_rot_3 <- create_complete_panel_indicator(event_study_df_all, "sample_monthly_rot_3", "time_index_monthly", 32)
complete_units_monthly_rot_4 <- create_complete_panel_indicator(event_study_df_all, "sample_monthly_rot_4", "time_index_monthly", 32)

# Identify units with complete panels for Monthly Rot 1 variants (different post-treatment periods)
complete_units_monthly_rot_1_post12 <- create_complete_panel_indicator(event_study_df_all, "sample_monthly_rot_1_post12", "time_index_monthly", 19)
complete_units_monthly_rot_1_post14 <- create_complete_panel_indicator(event_study_df_all, "sample_monthly_rot_1_post14", "time_index_monthly", 21)
complete_units_monthly_rot_1_post20 <- create_complete_panel_indicator(event_study_df_all, "sample_monthly_rot_1_post20", "time_index_monthly", 27)
complete_units_monthly_rot_1_post22 <- create_complete_panel_indicator(event_study_df_all, "sample_monthly_rot_1_post22", "time_index_monthly", 29)

# Identify units with complete panels for Monthly Rot 2 variants (different post-treatment periods)
complete_units_monthly_rot_2_post12 <- create_complete_panel_indicator(event_study_df_all, "sample_monthly_rot_2_post12", "time_index_monthly", 22)
complete_units_monthly_rot_2_post14 <- create_complete_panel_indicator(event_study_df_all, "sample_monthly_rot_2_post14", "time_index_monthly", 24)
complete_units_monthly_rot_2_post20 <- create_complete_panel_indicator(event_study_df_all, "sample_monthly_rot_2_post20", "time_index_monthly", 30)
complete_units_monthly_rot_2_post22 <- create_complete_panel_indicator(event_study_df_all, "sample_monthly_rot_2_post22", "time_index_monthly", 32)

# Update sample indicators to only include units with complete panels
event_study_df <- event_study_df_all %>%
  mutate(
    # Combined analysis sample indicators (with complete panel requirement)
    sample_combined_t4 = sample_combined_t4 & (id %in% complete_units_combined_t4),
    sample_combined_t5 = sample_combined_t5 & (id %in% complete_units_combined_t5),
    sample_combined_t6 = sample_combined_t6 & (id %in% complete_units_combined_t6),
    sample_combined_t7 = sample_combined_t7 & (id %in% complete_units_combined_t7),
    sample_combined_t8 = sample_combined_t8 & (id %in% complete_units_combined_t8),

    # Wave analysis sample indicators (with complete panel requirement)
    sample_wave_rot_1 = sample_wave_rot_1 & (id %in% complete_units_wave_rot_1),
    sample_wave_rot_2 = sample_wave_rot_2 & (id %in% complete_units_wave_rot_2),
    sample_wave_rot_3 = sample_wave_rot_3 & (id %in% complete_units_wave_rot_3),
    sample_wave_rot_4 = sample_wave_rot_4 & (id %in% complete_units_wave_rot_4),

    # Monthly analysis sample indicators (with complete panel requirement)
    sample_monthly_rot_1 = sample_monthly_rot_1 & (id %in% complete_units_monthly_rot_1),
    sample_monthly_rot_2 = sample_monthly_rot_2 & (id %in% complete_units_monthly_rot_2),
    sample_monthly_rot_3 = sample_monthly_rot_3 & (id %in% complete_units_monthly_rot_3),
    sample_monthly_rot_4 = sample_monthly_rot_4 & (id %in% complete_units_monthly_rot_4),

    # Monthly Rot 1 variants with different post-treatment periods (with complete panel requirement)
    sample_monthly_rot_1_post12 = sample_monthly_rot_1_post12 & (id %in% complete_units_monthly_rot_1_post12),
    sample_monthly_rot_1_post14 = sample_monthly_rot_1_post14 & (id %in% complete_units_monthly_rot_1_post14),
    sample_monthly_rot_1_post20 = sample_monthly_rot_1_post20 & (id %in% complete_units_monthly_rot_1_post20),
    sample_monthly_rot_1_post22 = sample_monthly_rot_1_post22 & (id %in% complete_units_monthly_rot_1_post22),

    # Monthly Rot 2 variants with different post-treatment periods (with complete panel requirement)
    sample_monthly_rot_2_post12 = sample_monthly_rot_2_post12 & (id %in% complete_units_monthly_rot_2_post12),
    sample_monthly_rot_2_post14 = sample_monthly_rot_2_post14 & (id %in% complete_units_monthly_rot_2_post14),
    sample_monthly_rot_2_post20 = sample_monthly_rot_2_post20 & (id %in% complete_units_monthly_rot_2_post20),
    sample_monthly_rot_2_post22 = sample_monthly_rot_2_post22 & (id %in% complete_units_monthly_rot_2_post22)
  )

# Print summary statistics
cat("\n=== Sample Sizes for Each Analysis Configuration ===\n\n")

cat("Monthly Analysis by Rotation (32 months, complete panels):\n")
cat(sprintf("  Rot 1: %d units\n", length(complete_units_monthly_rot_1)))
cat(sprintf("  Rot 2: %d units\n", length(complete_units_monthly_rot_2)))
cat(sprintf("  Rot 3: %d units\n", length(complete_units_monthly_rot_3)))
cat(sprintf("  Rot 4: %d units\n", length(complete_units_monthly_rot_4)))

cat("\nMonthly Rotation 1 with Different Post-Treatment Periods:\n")
cat(sprintf("  12mo post (19 periods): %d units\n", length(complete_units_monthly_rot_1_post12)))
cat(sprintf("  14mo post (21 periods): %d units\n", length(complete_units_monthly_rot_1_post14)))
cat(sprintf("  20mo post (27 periods): %d units\n", length(complete_units_monthly_rot_1_post20)))
cat(sprintf("  22mo post (29 periods): %d units\n", length(complete_units_monthly_rot_1_post22)))

cat("\nMonthly Rotation 2 with Different Post-Treatment Periods:\n")
cat(sprintf("  12mo post (22 periods): %d units\n", length(complete_units_monthly_rot_2_post12)))
cat(sprintf("  14mo post (24 periods): %d units\n", length(complete_units_monthly_rot_2_post14)))
cat(sprintf("  20mo post (30 periods): %d units\n", length(complete_units_monthly_rot_2_post20)))
cat(sprintf("  22mo post (32 periods): %d units\n", length(complete_units_monthly_rot_2_post22)))

cat("\nWave-based Analysis by Rotation (8 waves, complete panels):\n")
cat(sprintf("  Rot 1: %d units\n", length(complete_units_wave_rot_1)))
cat(sprintf("  Rot 2: %d units\n", length(complete_units_wave_rot_2)))
cat(sprintf("  Rot 3: %d units\n", length(complete_units_wave_rot_3)))
cat(sprintf("  Rot 4: %d units\n", length(complete_units_wave_rot_4)))

cat("\nCombined Analysis - All Rotations (Jan/May/Sep, complete panels):\n")
cat(sprintf("  t4 (4 periods): %d units\n", length(complete_units_combined_t4)))
cat(sprintf("  t5 (5 periods): %d units\n", length(complete_units_combined_t5)))
cat(sprintf("  t6 (6 periods): %d units\n", length(complete_units_combined_t6)))
cat(sprintf("  t7 (7 periods): %d units\n", length(complete_units_combined_t7)))
cat(sprintf("  t8 (8 periods): %d units\n", length(complete_units_combined_t8)))
```

```{r data-summary}
table(event_study_df$disab, event_study_df$disab_nwr)
```

Verify data structure for all three analysis types:
```{r verify-data-structure}
cat("\n=== Verifying Data Structure for All Analysis Types ===\n\n")

# Helper function to verify complete panel for a sample
verify_sample <- function(df, sample_col, time_col, expected_periods, sample_name) {
  sample_data <- df %>% filter(!!sym(sample_col))

  if (nrow(sample_data) == 0) {
    cat(sprintf("WARNING: %s has no data\n", sample_name))
    return(FALSE)
  }

  obs_per_unit <- sample_data %>%
    group_by(id) %>%
    summarize(n_obs = n_distinct(!!sym(time_col)), .groups = "drop")

  if (!all(obs_per_unit$n_obs == expected_periods)) {
    cat(sprintf("FAILED: %s - not all units have %d periods\n", sample_name, expected_periods))
    return(FALSE)
  }

  n_units <- nrow(obs_per_unit)
  cat(sprintf("PASSED: %s - %d units with %d periods each\n", sample_name, n_units, expected_periods))
  return(TRUE)
}

# Verify Combined Analysis samples
cat("Combined Analysis (Jan/May/Sep, all rotations):\n")
verify_sample(event_study_df, "sample_combined_t4", "time_index_combined", 4, "  Combined t4")
verify_sample(event_study_df, "sample_combined_t5", "time_index_combined", 5, "  Combined t5")
verify_sample(event_study_df, "sample_combined_t6", "time_index_combined", 6, "  Combined t6")
verify_sample(event_study_df, "sample_combined_t7", "time_index_combined", 7, "  Combined t7")
verify_sample(event_study_df, "sample_combined_t8", "time_index_combined", 8, "  Combined t8")

# Verify Wave Analysis samples
cat("\nWave-based Analysis by Rotation:\n")
verify_sample(event_study_df, "sample_wave_rot_1", "time_index_wave", 8, "  Wave Rot 1")
verify_sample(event_study_df, "sample_wave_rot_2", "time_index_wave", 8, "  Wave Rot 2")
verify_sample(event_study_df, "sample_wave_rot_3", "time_index_wave", 8, "  Wave Rot 3")
verify_sample(event_study_df, "sample_wave_rot_4", "time_index_wave", 8, "  Wave Rot 4")

# Verify Monthly Analysis samples
cat("\nMonthly Analysis by Rotation:\n")
verify_sample(event_study_df, "sample_monthly_rot_1", "time_index_monthly", 32, "  Monthly Rot 1")
verify_sample(event_study_df, "sample_monthly_rot_2", "time_index_monthly", 32, "  Monthly Rot 2")
verify_sample(event_study_df, "sample_monthly_rot_3", "time_index_monthly", 32, "  Monthly Rot 3")
verify_sample(event_study_df, "sample_monthly_rot_4", "time_index_monthly", 32, "  Monthly Rot 4")

# Verify time index columns have correct ranges
cat("\n=== Time Index Verification ===\n\n")

cat("time_index_monthly range by rotation:\n")
for (rot in 1:4) {
  rot_data <- event_study_df %>% filter(!!sym(paste0("rot_", rot)))
  if (nrow(rot_data) > 0) {
    cat(sprintf("  Rot %d: %d to %d\n", rot, min(rot_data$time_index_monthly), max(rot_data$time_index_monthly)))
  }
}

cat("\ntime_index_wave range (all rotations): ")
wave_data <- event_study_df %>% filter(sample_wave == 1)
cat(sprintf("%d to %d\n", min(wave_data$time_index_wave), max(wave_data$time_index_wave)))

cat("\ntime_index_combined range (Jan/May/Sep months): ")
combined_data <- event_study_df %>% filter(!is.na(time_index_combined))
cat(sprintf("%d to %d\n", min(combined_data$time_index_combined), max(combined_data$time_index_combined)))

# Verify calendar month mapping for combined analysis
cat("\nCalendar month mapping for Combined Analysis (time_index_combined):\n")
calendar_map <- event_study_df %>%
  filter(!is.na(time_index_combined)) %>%
  select(time_index_combined, calendar_year, calendar_month) %>%
  distinct() %>%
  arrange(time_index_combined)
print(head(calendar_map, 8))

# Verify treatment cohort columns
cat("\n=== Treatment Cohort Verification ===\n\n")
cat("treatment_cohort_monthly_disab values: ")
cat(paste(sort(unique(event_study_df$treatment_cohort_monthly_disab)), collapse = ", "), "\n")
cat("treatment_cohort_wave_disab values: ")
cat(paste(sort(unique(event_study_df$treatment_cohort_wave_disab)), collapse = ", "), "\n")
cat("treatment_cohort_combined_disab values: ")
cat(paste(sort(unique(event_study_df$treatment_cohort_combined_disab)), collapse = ", "), "\n")

# Verify rotation groups in data
cat("\n=== Rotation Group Verification ===\n\n")
rot_check <- event_study_df %>%
  select(rot) %>%
  distinct()
cat("Rotation groups in data:", paste(sort(rot_check$rot), collapse = ", "), "\n")
stopifnot(all(rot_check$rot %in% c(1, 2, 3, 4)))
cat("Rotation group verification passed: All 4 rotation groups present\n")
```

Save data:
```{r data-save}
# Create directory if it does not exist
if (!dir.exists(dirname(data_ada_path))) {
  dir.create(dirname(data_ada_path), recursive = TRUE)
}

# Select only columns relevant to empirical analysis
# Based on empirical_specs.csv requirements for all three analysis types
event_study_df <- event_study_df %>%
  select(
    # Core columns
    id, treated,

    # Time index columns for different analysis types
    time_index_monthly,    # Monthly analysis: relative to rotation (1 = first month, 7 = 7th month)
    time_index_wave,       # Wave analysis: interview_wave (1-8)
    time_index_combined,   # Combined analysis: Jan/May/Sep mapping (1-8)

    # Treatment cohort columns for different analysis types
    treatment_cohort_monthly_disab,    # Monthly: treatment at time_index = 7 (legacy, rotation 1 only)
    treatment_cohort_monthly_rot_1_disab,  # Monthly Rot 1: treatment at time_index = 7 (July 1990)
    treatment_cohort_monthly_rot_2_disab,  # Monthly Rot 2: treatment at time_index = 10 (July 1990)
    treatment_cohort_monthly_rot_3_disab,  # Monthly Rot 3: treatment at time_index = 9 (July 1990)
    treatment_cohort_monthly_rot_4_disab,  # Monthly Rot 4: treatment at time_index = 8 (July 1990)
    treatment_cohort_wave_disab,       # Wave: treatment at wave 3
    treatment_cohort_combined_disab,   # Combined: treatment at time_index = 3

    # Outcome variable
    labor_force_status,

    # Cluster variable
    state,

    # Sample indicators - Monthly by rotation (all 32 months)
    sample_monthly_rot_1, sample_monthly_rot_2, sample_monthly_rot_3, sample_monthly_rot_4,

    # Sample indicators - Monthly Rot 1 with different post-treatment periods
    sample_monthly_rot_1_post12, sample_monthly_rot_1_post14,
    sample_monthly_rot_1_post20, sample_monthly_rot_1_post22,

    # Sample indicators - Monthly Rot 2 with different post-treatment periods
    sample_monthly_rot_2_post12, sample_monthly_rot_2_post14,
    sample_monthly_rot_2_post20, sample_monthly_rot_2_post22,

    # Sample indicators - Wave by rotation (8 waves, first month of each wave)
    sample_wave_rot_1, sample_wave_rot_2, sample_wave_rot_3, sample_wave_rot_4,

    # Sample indicators - Combined (all rotations, Jan/May/Sep)
    sample_combined_t4, sample_combined_t5, sample_combined_t6, sample_combined_t7, sample_combined_t8,

    # Rotation indicators
    rot_1, rot_2, rot_3, rot_4,
    rot,

    # Auxiliary columns for debugging/verification
    calendar_year, calendar_month, interview_wave, sample_wave
  ) %>%
  # Convert id to numeric and state to numeric (required by did::att_gt)
  # state is haven_labelled from Stata import, needs conversion
  mutate(id = as.numeric(id),
         state = as.numeric(state)) %>%
  # Add legacy treatment_cohort column (required by apply_sample_restriction)
  # This will be overwritten by the cohort_varname specified in empirical_spec
  mutate(treatment_cohort = treatment_cohort_combined_disab)

print(paste("Columns in final dataset:", paste(colnames(event_study_df), collapse = ", ")))

# Print final dataset dimensions
cat(sprintf("\nFinal dataset: %d rows, %d columns\n", nrow(event_study_df), ncol(event_study_df)))

# Save
save(event_study_df, file = data_ada_path)
cat(sprintf("\nDataset saved to: %s\n", data_ada_path))
```
